dplyr Learning Notes
========================================================
by [Yimi Zhao](www.yimizhao.com)   
 
### Prepare the package and dataset 
```{r}
# install and load "dplyr" package
library(dplyr)
# load "CASchools" dataset in "AER" R package
library(AER)
data(CASchools)
# check the dataset
help(CASchools)
head(CASchools)
```

A few functions I used in dplyr
---------------------------------------
### tbl_df() - Create a data frame table 
```{r}
# user-friendly display the data
# A data frame tbl wraps a local data frame. The main advantage to using a tbl_df over a regular data
# frame is the printing: tbl objects only print a few rows and all the columns that fit on one screen,
# describing the rest of it as text.
CAS_tb<-tbl_df(CASchools)
# display more rows
print(CAS_tb,n=20)
```

### filter()- Return rows with matching conditions 
```{r}
# filter schools with more than 2000 students 
filter(CAS_tb, students >2000)
# filter schools in Kern and Los Angeles county
filter(CAS_tb, county=="Los Angeles" | county=="Kern")
filter(CAS_tb, county %in% c("Los Angeles","Kern"))
```
### select() - Select/rename variables by name 
```{r}
# select() keeps only the variables you mention; rename() keeps all variables
# to select school,computer, english, read, math variables
select(CAS_tb,school,computer, english, read,math)
# or 
select(CAS_tb,school,computer, english:math)
# special functions: start_with, ends_with, contains, matches, etc.
# seclect variables whose name starts/ends/contains/matches with "s"
select(CAS_tb, starts_with("s"))
select(CAS_tb, ends_with("s"))
select(CAS_tb, contains("s"))
select(CAS_tb, matches(".s."))
```
### Chaining or Pipelining 

```{r}
# write commands in a natural order by using the %>% operator (pronounced as “then”)
filter(select(CAS_tb,school,students,english:math),students>10000)
# can be write in this way
CAS_tb %>%
    select(school,students,english:math) %>%
    filter(students>10000)
```
### arrange() - Arrange rows by variables/columns

```{r}
arrange(CAS_tb,desc(students))
arrange(CAS_tb,county, school)
```
### mutate() - Add new variables/columns
```{r}
# add a variable refers to number of teachers per 100 students
head(mutate(CASchools,teacher_of_student = teachers/students*100))
#or 
CAS_tb %>%
    select(school,teachers, students) %>%
    mutate(teacher_of_student = teachers/students*100)
# or store the new variable 
CAS_tb <- CAS_tb %>%  mutate(teacher_of_student = teachers/students*100)
```
### group_by() and summarise()
```{r}
# summary the average students in each county
# using default R code
with(CASchools,tapply(students,county,mean))
aggregate(students~county,CASchools,mean)
# dplyr way
CASchools %>%
    group_by(county)%>%
    summarise(avg_students=mean(students))
```

### sample_n() and sample_frac - Sample n rows/fraction from a table 
```{r}
# similar with sample() and sample_int in base R
# sample 10 rows with/without replacement in base R way
CASchools[sample(nrow(CASchools),10),]
CASchools[sample(nrow(CASchools),10,replace=T),]
# dplyr way
sample_n(CASchools,10)
sample_n(CASchools,10,replace=T)
sample_frac(CASchools, 0.05) # sample 5% of the rows
```


